A family of accessibility measures derived from spatial interaction principles

### 2025-05-18 paper/manuscript.qmd
section 6

@@

In this section, we introduce a family of accessibility measures, using Wilson's spatial interaction modelling framework.

@@

Understanding $S_i$ as the inverse of $A_i$ offers a potential interpretation for the balancing factor, however, $S_i$ itself is not unclear. Also, the relationship in @eq-Ai-as-accessibility appears to be misaligned. In the preceding Section \@ref(grav-to-today), we noted that Hansen, and Stewart before him, defined accessibility as a partial sum of the demographic force $F$, essentially going from the demographic force ($F = G\frac{M_1\cdot M_2}{d_{12}^2}$ to the system-wide population potential (after losing $G$) $V_i = \sum_j \frac{M_j}{d_{ij}}$. While the inverse of Wilson's $A_i$ is pretty much identical to Hansen's accessibility, its functional role in Wilson's general model $T_{ij} = k W_i^{(1)} W_j^{(2)} f(c_{ij})$ is that of a balancing factor $k$ (maintaining proportionality).

@@

@@

To demonstrate the anatomy of $V_{ij}$ and $M_{ij}$ along with the partial sums of $V_i$ and $M_j$, Figure @fig-analytical-device-conc-accessibility demonstrates this work's accessibility analytical device for a simple 3 zone system. Each measure's most disaggregate value is $X_{ij}$, potential for spatial interaction from $i$ to $j$. The $X$ is a stand in for the $ij$ values of all the cases and their variants that will be described (i.e., $V_{ij}^0, M_{ij}^0, V_{ij}^T, M_{ij}^T, V_{ij}^S, M_{ij}^S$ and $V_{ij}^D, M_{ij}^D$). The single marginals represent the origin-side and destination-side weights of the zones. The total marginal represent the sum of a single marginal.

```{r}
#| label: fig-analytical-device-conc-accessibility
#| out-width: 70%
#| fig-cap: "The family of accessibility measures analytical device: labelling and associating ij flows, zonal weights, the single marginals, and the total marginal."

@@

As an additional overview, we define four distinct members of the constrained accessibility measure family, all delineated based on their constant $k$; either balancing factors $K^T$, $B_j$, $A_i$ depending on their case and variant:

>>>>

<!-- rafa 2: no need to keep all references in each citation block -->

@@

In this section, we introduce a family of accessibility measures inspired by Wilson's spatial interaction modelling framework.

@@

<!-- rafa 3: check sentence 'however, $S_i$ itself is not unclear' -->

<!-- rafa 4: check broken cross reference \@ref(grav-to-today) -->

Understanding $S_i$ as the inverse of $A_i$ offers a potential interpretation for the balancing factor, however, $S_i$ itself is not unclear. Also, the relationship in @eq-Ai-as-accessibility appears to be misaligned. In the preceding Section \@ref(grav-to-today), we noted that Hansen, and Stewart before him, defined accessibility as a partial sum of the demographic force $F$, essentially going from the demographic force ($F = G\frac{M_1\cdot M_2}{d_{12}^2}$) to the system-wide population potential (after losing $G$) $V_i = \sum_j \frac{M_j}{d_{ij}}$. While the inverse of Wilson's $A_i$ is pretty much identical to Hansen's accessibility, its functional role in Wilson's general model $T_{ij} = k W_i^{(1)} W_j^{(2)} f(c_{ij})$ is that of a balancing factor $k$ (maintaining proportionality).

@@

<!-- rafa 4: there is no $X$ in equation 23-->

@@

<!-- rafa 4: there is no $X$ in equation 23-->

To demonstrate the anatomy of $V_{ij}$ and $M_{ij}$ along with the partial sums of $V_i$ and $M_j$, Figure @fig-analytical-device-conc-accessibility illustrates the accessibility analytical framework we propose using a simple 3 zone system. Each measure's most disaggregate value is $X_{ij}$, potential for spatial interaction from $i$ to $j$. The $X$ is a stand in for the $ij$ values of all the cases and their variants that will be described (i.e., $V_{ij}^0, M_{ij}^0, V_{ij}^T, M_{ij}^T, V_{ij}^S, M_{ij}^S$ and $V_{ij}^D, M_{ij}^D$). The single marginals represent the origin-side and destination-side weights of the zones. The total marginal represent the sum of a single marginal.

```{r}
#| label: fig-analytical-device-conc-accessibility
#| out-width: 70%
#| fig-cap: "The family of accessibility measures analytical framework: labelling and associating ij flows, zonal weights, the single marginals, and the total marginal."

@@

As an additional overview, we define four distinct members of the constrained accessibility measure family, all delineated based on their constant $k$, which takes the form of either balancing factors $K^T$, $B_j$, $A_i$ depending on the indicator:
